---
title: "Calculating Validation Metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Valmetrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are several options/settings to calculate the validation metrics:   
**Minimum Allele 1 Probability Threshold** and **Maximum Allele 1 Probability Threshold**: The range of allele 1 probability thresholds for calculating the validation metrics, increasing in increments of 0.01 (i.e. if minimum is set to 0.5 and maximum is set to 1, will calculate metrics using a threshold of 0.5, 0.51, 0.52, 0.53, up to 1).   
**Minimum Allele 2 Probability Threshold** and **Maximum Allele 2 Probability Threshold**: The range of allele 2 probability thresholds for calculating the validation metrics, increasing in increments of 0.01 (i.e. if minimum is set to 0.5 and maximum is set to 1, will calculate metrics using a threshold of 0.5, 0.51, 0.52, 0.53, up to 1).   
**Major Contributor Sample ID**: The ID of the major contributor of the mixture. Once the reference folder is uploaded, this dropdown menu will auto-populate with the reference sample IDs.   
**Minor Contributor Sample ID**: The ID of the minor contributor of the mixture. Once the reference folder is uploaded, this dropdown menu will auto-populate with the reference sample IDs.    
**Remove SNPs If Missing Either Allele?**: If an allele 1 is inferred to be missing (reported as ```99```), that SNP will be automatically dropped from the final dataset. By default, if an allele 2 is inferred to be missing and the allele 2 probability is above the allele 2 probability threshold, the SNP is reported as homozygous for allele 1. However, selecting this option will result in dropping the SNP if the allele 2 probability of the missing allele 2 is above the allele 2 probability threshold, instead of reporting the SNP as homozygous for allele 1.   

If calculating validation metrics, reference genotypes are required to calculate genotype accuracy. MixDerR will calculate the metrics for the range of allele 1 probability thresholds and allele 2 probability thresholds specified by the user. The final output file looks as such:

```{r table6, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl = "
| A1 cutoff | A2 cutoff    | Total SNPs | N No Ref | N SNPs tested | N Genotypes Correct | Genotype Accuracy | Heterozygosity |
|:-----------:|:---------:|:--------:|:--------:|:--------:|:--------:|:--------:|:--------:|
| 0.99    | 0.01 | 9735 | 8 | 9727 | 9549 | 0.9817 | 0.456 |
| 0.99    | 0.02 | 9735 | 8 | 9727 | 9548 | 0.9815 | 0.456 |
| 0.99    | 0.03 | 9735 | 8 | 9727 | 9548 | 0.9815 | 0.456 |
"
cat(tabl)
```
